Energy Dependence of Synchrotron X-Ray Rims in Tycho's Supernova Remnant

### 2014-12-23 manuscript/paper-tycho.tex
Finish telecon todos, update ms tables/plots
Address remaining telecon issues, minor ipynb refactoring.
Spectral variation, eyeball fits compared to width-energy.
Tables for eyeball fits, damped srcutlog fits.
New manuscript fig/table layout. Discard old figs/tables.
Add CASA tutorial notes/etc from Nov 2014.

% e- cut-off energy result
With a diffusion coefficient in hand, we may determine the electron spectrum
cut-off energy by equating synchrotron loss and diffusive acceleration
timescales.  For low energies and small synchrotron losses (cooling time longer
than acceleration time), electrons may be efficiently accelerated; near or
above the cut-off energy, electrons will radiate or escape too rapidly to be
accelerated to higher energies and the energy spectrum drops off steeply.  The
cut-off energy is given as:
\begin{align} \label{eq:ecut}
    \Ecut =
        &\left(8.3\unit{TeV}\right)^{2/(1+\mu)}
        \left(\frac{B_0}{100 \muG}\right)^{-1/(1+\mu)} \nonumber \\
        &\times \left(\frac{v_s}{10^8 \unit{cm\;s^{-1}}}\right)^{2/(1+\mu)}
        \eta^{-1 / (1+\mu)} .
\end{align}
Here $B_0$ is the magnetic field immediately downstream of the shock, i.e.
$B_0 = B(x=0)$.
This result is derived by \citet{parizot2006} for $\mu=1$ assuming a strong
shock with compression ratio 4 and isotropic magnetic turbulence on both sides
of the shock.

@@

@@

@@

% Example loss-limited fits
\begin{table}
    \scriptsize
    \centering
    \caption{Loss-limited fits for Regions 1, 16
    \label{tab:fits-loss}}
    \input{tab-fits-loss.tex}
    \tablecomments{Model fits computed at several fixed values of $\mu$ with
    $\eta_2$, $B_0$ free and constant magnetic field $B(x) = B_0$.
    Region 1 has 2 degrees of freedom (dofs); Region 16, 3.}
\end{table}

% Example damping fits
\begin{table}
    \scriptsize
    \centering
    \caption{Magnetic damping fits for Regions 1, 16
    \label{tab:fits-damp}}
    \input{tab-fits-damp.tex}
    \tablecomments{Model fits computed at several fixed values of $a_b$ with
    $\eta_2$, $B_0$ free and damped magnetic field (equation~\eqref{eq:bdamp}).
    Fit degrees of freedom are as in Table~\ref{tab:fits-loss}}
\end{table}

@@

% As above but with eta2=1 fixed!
\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits for all regions, $\mu = \eta_2 = 1$
    \label{tab:fits-all-eta2one}}
    \input{tab-fits-all-eta2one.tex}
    \tablecomments{Best damped fits select best $a_b$ value below $0.01$ from
    the range shown in Table~\ref{tab:fits-damp}, enforcing moderate damping.
    Fits for Regions 1--3, 5, 11, 12, 18, and 19 have 3 degrees of freedom; all
    others have 4.  The manual selection of a best $a_b$ value may be construed
    as removing one additional dof.}
\end{table}

% TODO EXPLAIN/DISCUSS

% As above but with eta2=1 fixed!
\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits with $\Bmin = 0$ and $\mu = \eta_2 = 1$
    \label{tab:fits-all-bminzero}}

    \begin{tabular}{@{} l rr rrr @{}}
    \toprule
    {} & \multicolumn{2}{c}{Loss-limited}
       & \multicolumn{3}{c}{Damped, $a_b \leq 0.01$} \\
    \cmidrule(lr){2-3} \cmidrule(l){4-6}
    Region & $B_0$ ($\mu$G) & $\chi^2$
           & $B_0$ ($\mu$G) & $\chi^2$ & $a_b$ \\
    \midrule
    1 & 183 & 35.63 & 2.43 & 629.61 & 0.010 \\
    2 & 313 & 80.81 & 239 & 100.66 & 0.010 \\
    3 & 427 & 21.48 & 405 & 21.70 & 0.010 \\
    \cmidrule{1-6}
    4 & 285 & 51.30 & 174 & 43.82 & 0.010 \\
    5 & 289 & 19.30 & 196 & 37.33 & 0.010 \\
    6 & 411 & 174.32 & 90.2 & 61.50 & 0.005 \\
    7 & 419 & 51.30 & 277 & 11.13 & 0.006 \\
    8 & 388 & 48.06 & 205 & 9.92 & 0.006 \\
    9 & 415 & 65.45 & 143 & 14.01 & 0.005 \\
    10 & 466 & 17.28 & 307 & 1.60 & 0.005 \\
    \cmidrule{1-6}
    11 & 356 & 11.95 & 313 & 16.08 & 0.010 \\
    12 & 318 & 9.60 & 258 & 15.45 & 0.010 \\
    13 & 400 & 53.11 & 240 & 10.00 & 0.006 \\
    \cmidrule{1-6}
    14 & 384 & 103.43 & 52.3 & 9.90 & 0.005 \\
    15 & 432 & 70.66 & 255 & 20.18 & 0.005 \\
    16 & 493 & 15.13 & 430 & 3.88 & 0.006 \\
    17 & 468 & 61.84 & 194 & 29.05 & 0.004 \\
    \cmidrule{1-6}
    18 & 283 & 27.87 & 206 & 71.82 & 0.010 \\
    19 & 401 & 37.07 & 144 & 8.14 & 0.005 \\
    20 & 464 & 122.00 & 140 & 94.37 & 0.004 \\
    \bottomrule
    \end{tabular}

    \tablecomments{See comments for Table~\ref{tab:fits-all-eta2one}.}
\end{table}

@@

@@

@@

@@

To measure synchrotron cut-offs in Tycho, we fitted ``rim'' spectra
(Section~\ref{sec:spec}) to the XSPEC model \texttt{srcut}, modified to fit in
log-frequency units; we fix the radio spectral index $\alpha = 0.58$
\citep{sun2011} as done in our profile modeling and fit for column density
$N_{\mt{H}}$ and cut-off frequency $\nu_{\mt{cut}}$ in each region.
Table~\ref{tab:fits-srcutlog} lists best spectrum fit parameters for each
region.  The fitted cut-off frequency is typically $0.3 \unit{keV/h}$,
consistent with fits by \citet{hwang2002}, but Regions 11, 12 have unusually
high cut-off frequencies $0.55$, $0.88 \unit{keV/h}$ consistent with harder rim
spectra (Table~\ref{tab:spec}).

@@

\begin{table}
    \scriptsize
    \centering
    \caption{Synchrotron roll-off, $\eta_2$, and fits for $\mu = 1$
    \label{tab:fits-srcutlog}}
    \input{tab-srcutlog.tex}
    \tablecomments{\texttt{srcut} fits performed in log-frequency coordinates;
        here $h$ is Planck's constant. $\eta_2$ values are computed from
        equation~\eqref{eq:cutoff} and held fixed in model fits.}
\end{table}

@@

% TODO We have NOT considered the relative intensities of X-ray / radio rims!

@@

% ================================
% Best fit information for SN 1006
% ================================
\section{Best fits, SN 1006}

@@

\clearpage
\begin{figure}
    \centering
    \iftoggle{manuscript}{
        \includegraphics[width=0.48\textwidth]{figures/mE-damp-sn1006.pdf}
    }{
        \includegraphics[width=0.35\textwidth]{figures/mE-damp-sn1006.pdf}
    }
    \caption{Curves of $\mE$ values for best damping fits for all regions,
        SN 1006}
\end{figure}

@@

% ========================================
% Best Tycho fits with Bmin = 2 microGauss
% ========================================

%\clearpage
%\section{Best fits, Tycho, with $\Bmin=2\muG$}
%
%\begin{table}[ht]
%    \scriptsize
%    \centering
%    \caption{Best model fits for all regions, $\mu = 1$, $\Bmin=2\muG$}
%    \input{suppl-fits-all-bmin2.tex}
%\end{table}
%
%% As above but with eta2=1 fixed!
%\begin{table}[ht]
%    \scriptsize
%    \centering
%    \caption{Best model fits for all regions, $\mu = \eta_2 = 1$,
%        $\Bmin=2\muG$}
%    \input{suppl-fits-all-bmin2-eta2one.tex}
%\end{table}
%
%% Only show these giant tables in 2col... no need to print these
%\iftoggle{manuscript}{
%}{
%    % ===============================
%    % Damping best fit sea of numbers
%    % ===============================
%    \clearpage
%    \section{Magnetic damping best fit tables}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to individual regions, Filament 1, $\mu = 1$}
%        \input{suppl-fits-damp-f1.tex}
%    \end{table*}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to individual regions, Filament 2, $\mu = 1$}
%        \input{suppl-fits-damp-f2.tex}
%    \end{table*}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to individual regions, Filament 2 (cont.), $\mu = 1$}
%        \input{suppl-fits-damp-f2-pt2.tex}
%    \end{table*}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to individual regions, Filament 3, $\mu = 1$}
%        \input{suppl-fits-damp-f3.tex}
%    \end{table*}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to individual regions, Filament 4, $\mu = 1$}
%        \input{suppl-fits-damp-f4.tex}
%    \end{table*}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to individual regions, Filament 5, $\mu = 1$}
%        \input{suppl-fits-damp-f5.tex}
%    \end{table*}
%
%    % ================================
%    % Damping best fit tables, SN 1006
%    % ================================
%    \clearpage
%    \section{Magnetic damping fit tables, SN 1006}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to SN1006 filaments, $\mu = 1$}
%        \input{suppl-fits-damp-sn1006.tex}
%    \end{table*}
%
%    \clearpage
%    \section{Magnetic damping fit tables, SN 1006, no e- cutoff}
%
%    \begin{table*}[ht]
%        \scriptsize
%        \centering
%        \caption{Damped fits to SN1006 filaments, $\mu = 1$, no e- cutoff}
%        \input{suppl-fits-damp-nocut-sn1006.tex}
%    \end{table*}
%}

>>>>

% e- cut-off energy result
With a diffusion coefficient in hand, we may determine the electron spectrum
cut-off energy by equating synchrotron loss and diffusive acceleration
timescales.  For low energies and small synchrotron losses (cooling time longer
than acceleration time), electrons may be efficiently accelerated; near or
above the cut-off energy, electrons will radiate or escape too rapidly to be
accelerated to higher energies and the energy spectrum drops off steeply.  The
cut-off energy is given as:
\begin{align} \label{eq:ecut}
    \Ecut =
        &\left(8.3\unit{TeV}\right)^{2/(1+\mu)}
        \left(\frac{B_0}{100 \muG}\right)^{-1/(1+\mu)} \nonumber \\
        &\times \left(\frac{v_s}{10^8 \unit{cm\;s^{-1}}}\right)^{2/(1+\mu)}
        \eta^{-1 / (1+\mu)} .
\end{align}
Here $B_0$ is the magnetic field immediately downstream of the shock, i.e.
$B_0 = B(x=0)$; our notation departs somewhat from typical notation in denoting
downstream field by $B_0$.
This result is derived by \citet{parizot2006} for $\mu=1$ assuming a strong
shock with compression ratio 4 and isotropic magnetic turbulence on both sides
of the shock.

@@

% TODO discuss srcutlog fitting (move text up from below)
\begin{table}
    \scriptsize
    \centering
    \caption{Synchrotron roll-off, $\eta_2$, and fits for $\mu = 1$
    \label{tab:spec-srcutlog}}
    \input{tab-srcutlog.tex}
    \tablecomments{\texttt{srcut} fits performed in log-frequency coordinates;
        here $h$ is Planck's constant. $\eta_2$ values are computed from
        equation~\eqref{eq:cutoff} and held fixed in model fits.}
\end{table}

@@

% TODO cull/rewrite all references to tab:fits-loss, tab:fits-damp
% Since tables are removed, adjust discussion of eta2, mu effects on fits.
% and effects of varying ab value?

@@

% Example loss-limited fits
%\begin{table}
%    \scriptsize
%    \centering
%    \caption{Loss-limited fits for Regions 1, 16
%    \label{tab:fits-loss}}
%    \input{tab-fits-loss.tex}
%    \tablecomments{Model fits computed at several fixed values of $\mu$ with
%    $\eta_2$, $B_0$ free and constant magnetic field $B(x) = B_0$.
%    Region 1 has 2 degrees of freedom (dofs); Region 16, 3.}
%\end{table}
%
%% Example damping fits
%\begin{table}
%    \scriptsize
%    \centering
%    \caption{Magnetic damping fits for Regions 1, 16
%    \label{tab:fits-damp}}
%    \input{tab-fits-damp.tex}
%    \tablecomments{Model fits computed at several fixed values of $a_b$ with
%    $\eta_2$, $B_0$ free and damped magnetic field (equation~\eqref{eq:bdamp}).
%    Fit degrees of freedom are as in Table~\ref{tab:fits-loss}}
%\end{table}

@@

\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits for all regions, $\mu = \eta_2 = 1$
    \label{tab:fits-all-eta2one}}
    \input{tab-fits-all-eta2one.tex}
    \tablecomments{Fits for Regions 1--3, 5, 11, 12, 18, and 19 have 3 degrees
    of freedom; all others have 4.  The choice of a best $a_b$ value may be
    construed as removing one additional dof.}
\end{table}

% Integrate srcutlog into textual discussion (!)

\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits for all regions, $\eta_2$ derived from
        \texttt{srcut} fits
    \label{tab:fits-all-srcutlog}}
    \input{tab-fits-all-srcutlog.tex}
    \tablecomments{$\eta_2$ values are computed from equation~\eqref{eq:cutoff}
    and held fixed in model fits.  See comments for
    Table~\ref{tab:fits-all-eta2one} as well.}
\end{table}

@@

% TODO discuss this a bit further.  Explain that it doesn't matter for us since
% we are not exploring eta2 dependence of fits etc, but something useful to
% note

@@

\subsection{Lengthscale analysis}

% TODO Short blurb

\begin{figure}
    \centering
    \plotone{figures/lengthscales_fwhm.pdf} \\
    \caption{Stuff
        \label{fig:length}}
\end{figure}

\begin{figure}
    \centering
    \plotone{figures/lengthscales_fwhm_dimless.pdf} \\
    \caption{Stuff.  Dashed line plots scaled FWHM = $4.6$, the projection
        factor for a spherical shell (CITE)  % TODO
        \label{fig:length-dimless}}
\end{figure}

\begin{table*}
    \scriptsize
    \centering
    \caption{Lengthscale analysis
        \label{tab:lengthscales}}
    \input{tab-lengthscale.tex}
    \tablecomments{All derived from fits with $\eta_2 = 1$ and such.  Ratio of
        $l_{\mt{ad}}/l_{\mt{diff}}$ is same for loss-limited and damped fits
    (see text).
    All values computed at fiducial energy $2 \unit{keV}$.}
\end{table*}

@@

% TODO move this up in discussion to EARLIER.

@@

To measure synchrotron cut-offs in Tycho, we fitted ``rim'' spectra
(Section~\ref{sec:spec}) to the XSPEC model \texttt{srcut}, modified to fit in
log-frequency units; we fix the radio spectral index $\alpha = 0.58$
\citep{sun2011} as done in our profile modeling and fit for column density
$N_{\mt{H}}$ and cut-off frequency $\nu_{\mt{cut}}$ in each region.
Table~\ref{tab:fits-all-srcutlog} lists best spectrum fit parameters for each
region.  The fitted cut-off frequency is typically $0.3 \unit{keV/h}$,
consistent with fits by \citet{hwang2002}, but Regions 11, 12 have unusually
high cut-off frequencies $0.55$, $0.88 \unit{keV/h}$ consistent with harder rim
spectra (Table~\ref{tab:spec}).

@@

@@

\begin{table}
    \scriptsize
    \centering
    \caption{Eyeballed fits
        \label{tab:fits-eyeball}}
    \input{tab-fits-eyeball.tex}
    \tablecomments{$\infty$ indicates that we favor a loss-limited fit here
    ($a_b > 10$\% of shock radius $r_s$)}
\end{table}

@@

% ======================================
% Extra tables/fit information for Tycho
% ======================================
\section{Extra Tycho results}

% TODO EXPLAIN/DISCUSS BMIN=ZERO INLINE (not here obviously)
\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits with $\Bmin = 0$ and $\mu = \eta_2 = 1$}
    \input{suppl-fits-all-bmin0-eta2one.tex}
    \tablecomments{See comments for Table~\ref{tab:fits-all-eta2one}.}
\end{table}

% TODO EXPLAIN/DISCUSS ETA2 INLINE
\begin{table}
    \scriptsize
    \centering
    \caption{Best model fits with $\eta_2 = 10$ fixed}
    \input{suppl-fits-all-eta2ten.tex}
    \tablecomments{See comments for Table~\ref{tab:fits-all-eta2one}.}
\end{table}

% TODO EXPLAIN
\begin{figure*}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/specvar-dgamma.pdf}
    \caption{Spectral variation as a function of energy for damped best fit
    parameters, w/ different choice of $\eta_2$.}
\end{figure*}

% ================================
% Best fit information for SN 1006
% ================================
\clearpage
\section{Best fits, SN 1006}

@@

@@
